A simple model to describe how outflows and rotation shape the Lyman-$\alpha$ emission line in galaxies

### 2017-03-02 paper/paper.tex
title, intro

\title{Influence of galaxy rotation and outflows in the Lyman-$\alpha$
  spectral line}
\shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}

@@

In the early Universe the most abundant elements were Hydrogen(H) and
Helium(He), causing young galaxies to be rich in H atoms. Due to
stellar activity in a galaxy, the surrounding gas is ionized, so H
atoms excite and emit radiation at definite frequencies. The most
basic of these frequencies corresponds to the \lya line equal to $2.47
\times 10^{15}\;Hz$ \cite{PartridgePeebles}. This is equivalent to a
wavelength of 1215.67 $\AA$ cause by the transition of the H atom
electron from levels $2\rightarrow1$ dubbed Lyman Alpha ( \lya)
transition, discovered by Theodore Lyman in 1906. \cite{LymanBio} \\

Due to the amount of H atoms inside a galaxy, the whole body becomes a
strong \lya radiator. If the galaxy spectra shows a strong \lya line,
it classifies as a Lyman Alpha Emitter
(LAE). (\cite{DjorgovskiThompson}, \cite{Rhoads00},
\cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
\cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12},
\cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013},
\cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
\cite{Fumagalli2015}). However, when a \lya photon is emitted inside
the galaxy, it travels through its interstellar medium (ISM). During
the photon's path it can be absorbed and re-emitted by other H
atoms. The new frequency of the photon is different that the initial
one, in an observer frame of reference due to the atom's velocity. In
LAEs, the state of the ISM gas, before and after a photon's
re-emission is pretty much the same. This allows a random walk
approximation and lets us consider photon-atom encounters as
scatterings. A photon's scattering can happen several times and stops
only until the photon is able to escape the galaxy. \\

In a static galaxy, this random walk process produces a spectrum with
two equal and symmetric peaks around the natural \lya wavelength and
zero intensity at the center (Fig. A3 from \cite{CLARA}). If now the
gas has a bulk velocity, the shape of the \lya profile changes. We
explore these effects in this paper by proposing a new model for a
LAE. The model consists of a spherical distribution of H atoms
undergoing a solid body rotation and radial expansion (outflows). \\

>>>>

\title{A simple model to describe how outflows and rotation shape the Lyman-$\alpha$
  emission line in galaxies}
\shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}

@@

Currently LAEs are commonly targetted in wide area galaxy surveys.
They have been effectively used to study galaxy evolution,
cosmology and the thermal history of the Universe.
This has been able through the study of their spatial distribution and
their global line shape.
The recent release of MUSE has caused an switch in the kind of studies
that can be performed on LAEs.

Thanks to MUSE's ability to spatially resolve a galaxy, we are now able
to infer detailed kinematic maps for nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

On the ISM's features that plays an important role in shaping the \lya
is HI kinematics. 
In a static HI medium the \lya line has  two equal and symmetric peaks
around the natural \lya wavelength and zero intensity at the line's
center.
If the ISM has an outflow velocity profile, the line becomes
asymmetrical with a more pronounced read peak.
If the galaxy rotates, the line shows different amounts of Doppler
shifts that would also modify the profile.
gas has a bulk velocity, the shape of the \lya profile changes. We
explore these effects in this paper by proposing a new model for a
LAE. The model consists of a spherical distribution of H atoms
undergoing a solid body rotation and radial expansion (outflows). \\

\cite{DjorgovskiThompson}, \cite{Rhoads00},
\cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
\cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12},
\cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013},
\cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
\cite{Fumagalli2015}.
